---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `ClinTrialX`

<!-- badges: start -->
[![R-CMD-check](https://github.com/ineelhere/clintrialx/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ineelhere/clintrialx/actions/workflows/R-CMD-check.yaml)
[![License: Apache-2.0](https://img.shields.io/badge/license-Apache--2.0-blue.svg)](https://opensource.org/licenses/Apache-2.0)
[![pkgdown](https://img.shields.io/badge/pkgdown-docs-blue.svg)](https://ineelhere.github.io/clintrialx/)

<!-- badges: end -->


The goal of `{clintrialx}` is to fetch clinical trials data from freely available registries. Currently, it supports the ClinicalTrials.gov registry using its [V2 API](https://clinicaltrials.gov/data-api/api).

## Installation

You can install this package from [GitHub](https://github.com/ineelhere/clintrialx) with:

```r
# install.packages("devtools")
devtools::install_github("ineelhere/clintrialx")
```
### Check installation
```r
library(clintrialx)
```

## Query Data based on NCT IDs

Fetch one or multiple trial records based on NCT IDs. You can opt to fetch some specific fields or all fields available at source (default).

```{r example1}
library(clintrialx)
ctg_get_nct(c("NCT02967965", "NCT04000165", "NCT01007279", "NCT02376244", "NCT01179776"),
                          fields = c("NCT Number", "Study Title", "Study Status", "Sponsor"))


```

## Query Data based on fileds

Supports filtering by condition, location, title keywords, intervention, and overall status.


```{r example2}
ctg_get_fields(
     condition = "Cancer",
     location = "Kolkata",
     title = NULL,
     intervention = "Drug",
     status = c("ACTIVE_NOT_RECRUITING", "RECRUITING"),
     page_size = 10
)

```


## Data Sources

You can fetch version information directly from the package:

```{r example3}
version_info(source = "clinicaltrials.gov")


```

## Get Involved

🚀 **Ready to contribute?** We welcome contributions to make `clintrialx` even better. Check out our [contributing guidelines](https://github.com/ineelhere/clintrialx/blob/main/CONTRIBUTING.md) to get started.

💬 **Questions or Feedback?** Feel free to open an issue on our [GitHub Issues page](https://github.com/ineelhere/clintrialx/issues).

🌟 **Enjoying `clintrialx`?** Please consider giving us a star on [GitHub](https://github.com/ineelhere/clintrialx)! Your support helps us grow and improve.

More updates to come. Happy coding! 🎉
