---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# clintrialx

<!-- badges: start -->
[![R-CMD-check](https://github.com/ineelhere/clintrialx/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ineelhere/clintrialx/actions/workflows/R-CMD-check.yaml)
[![License: Apache-2.0](https://img.shields.io/badge/license-Apache--2.0-blue.svg)](https://opensource.org/licenses/Apache-2.0)
[![pkgdown](https://img.shields.io/badge/pkgdown-docs-blue.svg)](https://ineelhere.github.io/clintrialx/)
<!-- badges: end -->

The goal of `{clintrialx}` is to fetch clinical trials data from freely available registries. Currently, it supports the ClinicalTrials.gov registry using its [V2 API](https://clinicaltrials.gov/data-api/api).

## Installation

You can install this package from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("ineelhere/clintrialx")
```

## Example

This is a basic example that shows how to download data based on NCT ID(s):

```{r example1}
library(clintrialx)
fetch_specific_trial_data("NCT04000165")
```

Fetch only a few/specific fields:

```{r example2}
library(clintrialx)
fetch_specific_trial_data("NCT04000165", fields = c("NCT Number", "Study Title", "Study Status", "Sponsor"))
```

You can also download data for multiple NCT IDs:

```{r example3}
library(clintrialx)
fetch_specific_trial_data(c("NCT02967965", "NCT04000165", "NCT01007279", "NCT02376244", "NCT01179776"))
```

Similarly, you can query only desired fields:

```{r example4}
library(clintrialx)
fetch_specific_trial_data(c("NCT02967965", "NCT04000165", "NCT01007279", "NCT02376244", "NCT01179776"),
                          fields = c("NCT Number", "Study Title", "Study Status", "Sponsor"))
```

## Data Sources

You can fetch version information directly from the package:

```{r example}
library(clintrialx)
version_info(source = "clinicaltrials.gov")
```

## Get Involved

🚀 **Ready to contribute?** We welcome contributions to make `clintrialx` even better. Check out [contributing guidelines](https://github.com/ineelhere/clintrialx/blob/main/CONTRIBUTING.md) to get started.

💬 **Questions or Feedback?** Feel free to open an issue on [GitHub Issues page](https://github.com/ineelhere/clintrialx/issues).

🌟 **Enjoying `clintrialx`?** Please consider giving a star on [GitHub](https://github.com/ineelhere/clintrialx)! Your support helps this project grow and improve.

More updates to come. Happy coding! 🎉
